Artificial Ecosystem Selection

### 2019-09-18 R/fluxes.Rmd
Update protocol

library(tidyverse)
library(lubridate)
library(broom)
library(broman)
library(purrr)
library(kableExtra)

@@

top_jars_gen1 <- c(2, 3, 13, 16)

@@

n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
```

@@

A set of `r n1` replicate ecosystems were generated
with 50 g of soil (sterile potting mix plus 5% inoculum by mass)
and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
These jars were initially kept at atmospheric concentrations of methane, but did not 
display significant methane oxidation rates after ~6 weeks. As a result, 
they were capped and spiked to
a headspace concentration of approximately 1000 ppm CH4. Both CH4 and CO2 were monitored
daily and it was determined that flushing and respiking was required every three
days in order to maintain aerobic conditions and an elevated CH4 concentration. After about 8 weeks
of incubation (2 weeks at 1000 ppm), we finally observed significant (negative)
methane fluxes. From the `r n1` jars, we selected the top 4 
(jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.

For the first transfer, we used 10% inoculum intead of 5% inoculum and generated
`r n2` jars. After one week we observed
significant, negative methane fluxes that exceeded the first set of ecosystems. This
may have been a response to selection, a biomass effect due to the larger inoculum,
or both. After two weeks, we observed dampened fluxes compared to week one, but still
greater flux than before the first transfer. The elevated fluxes in week 1 compared
to week 2 could have been a result of using two different standard curves for t0 and t1
and for t2 and t3 in week 1 or due to some temporal effect during the incubations 
(e.g. response to water addition or initial available nutrients or growth).
In order to truly demonstrate a response to selection we will generate at
least one more set of ecosystems through a transfer with the same inoculum size
(10%).

@@

conc$selected[conc$JAR %in% top_jars_gen1 & conc$transfer == '1'] <- T
conc$selected[!(conc$JAR %in% top_jars_gen1) | conc$transfer != '1'] <- F
```

@@

flux <- 
flux %>%
  mutate(tidy_lm = map(fit_lm, tidy),
         tidy_glm = map(fit_glm, tidy)) %>% 
unnest(tidy_lm, tidy_glm) %>% 
filter(term == 'days') %>% 
select(transfer, Date, JAR, selected, lm_umold = estimate, glm_umold = estimate1)

all_par <- flux$lm_umold[flux$Date == '20190527']
offspr1 <- flux$lm_umold[flux$Date == '20190607']
offspr2 <- flux$lm_umold[flux$Date == '20190614']
sel_par <- flux$lm_umold[flux$transfer == '1' & flux$selected == T]

all_par_log <- flux$glm_umold[flux$Date == '20190527']
offspr1_log <- flux$glm_umold[flux$Date == '20190607']
offspr2_log <- flux$glm_umold[flux$Date == '20190614']
sel_par_log <- flux$glm_umold[flux$transfer == '1' & flux$selected == T]

@@

ggplot(flux, aes(x = lm_umold)) +
  geom_histogram(aes(fill = interaction(selected, Date))) +
  facet_grid(Date ~ .) +
  scale_fill_manual(values = c('black', 'red', 'blue', 'darkblue')) +
  geom_vline(xintercept = mean(all_par)) +
  geom_vline(xintercept = mean(offspr1), color = 'blue') +
  geom_vline(xintercept = mean(offspr2), color = 'darkblue') +
  geom_vline(xintercept = mean(sel_par), color = 'red') +
  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "LM Fluxes")

ggplot(flux, aes(x = Date, y = lm_umold)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = interaction(selected, Date))) +
  scale_color_manual(values = c('black', 'red', 'blue', 'darkblue')) + 
  geom_hline(yintercept = mean(all_par)) +
  geom_hline(yintercept = mean(offspr1), color = 'blue') +
  geom_hline(yintercept = mean(offspr2), color = 'darkblue') +
  geom_hline(yintercept = mean(sel_par), color = 'red')  + 
  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "LM Fluxes")

ggplot(flux, aes(x = glm_umold)) +
  geom_histogram(aes(fill = interaction(selected, Date))) +
  facet_grid(Date ~ .) +
  scale_fill_manual(values = c('black', 'red', 'blue', 'darkblue')) +
  geom_vline(xintercept = mean(all_par_log)) +
  geom_vline(xintercept = mean(offspr1_log), color = 'blue') +
  geom_vline(xintercept = mean(offspr2_log), color = 'darkblue') +
  geom_vline(xintercept = mean(sel_par_log), color = 'red') +
  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "GLM Fluxes")

ggplot(flux, aes(x = Date, y = glm_umold)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = interaction(selected, Date))) +
  scale_color_manual(values = c('black', 'red', 'blue', 'darkblue')) + 
  geom_hline(yintercept = mean(all_par_log)) +
  geom_hline(yintercept = mean(offspr1_log), color = 'blue') +
  geom_hline(yintercept = mean(offspr2_log), color = 'darkblue') +
  geom_hline(yintercept = mean(sel_par_log), color = 'red') +
  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "GLM Fluxes")
```

@@

nsh1 <- ns_heritability(all_par, offspr1, sel_par) 
nsh2 <- ns_heritability(all_par, offspr2, sel_par) 
nsh1_log <- ns_heritability(all_par_log, offspr1_log, sel_par_log)
nsh2_log <- ns_heritability(all_par_log, offspr2_log, sel_par_log)
```

Based on the mean ecosystem-level trait values from the first and second set
of ecosystems (week 2), we calculated narrow-sense heritability as $h^2 = R / S$, which
was `r myround(nsh2)` for the linear model and `r myround(nsh2_log)` for the exponential
model. Granted, this is also not a valid estimate, because of the difference
in inoculum size between transfer 1 and 2. We will calculate this again between
transfer 2 and 3 to get a more accurate estimate.

@@

```{r identify_top_jars} 
flux %>% 
  group_by(Date) %>% 
  top_n(-4, lm_umold) %>% 
  kable(caption = "Linear model fluxes. transfer is the ecosystem generation,
        Date is the flux measurement date, JAR lists the top 4 jars from that sample
        date in terms of methane flux from a linear model, 
        selected indicates whether they were the
        selected parents for transfer 1, lm_umold is uMol/d flux based on a linear
        regression and glm_umold is uMol/d flux based on a glm with family = Gaussian
        and link = log") %>% 
  kable_styling()
flux %>% 
  group_by(Date) %>% 
  top_n(-4, glm_umold) %>% 
  kable(caption = "Generalized linear model link = log fluxes. Same as above except
        JAR indicates top 4 jars from a glm.") %>% 
  kable_styling()
```

>>>>

library(tidyverse)
library(lubridate)
library(broom)
library(broman)
library(purrr)
library(kableExtra)
library(morris)

@@

top_jars_pas1 <- c(2, 3, 13, 16)
top_jars_pas2 <- c(6, 8, 17, 21)

@@

n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))

# The two SCs from passage 2, week 1
sc_1 <- function(x) {(0.1904*x)+58.823}
sc_2 <- function(x) {(0.1427*x)+99.075}

# Convert the saved concentrations (using two different curves) to using
# one or the other
conc <- 
  conc  %>% 
  mutate_cond(Date == "20190607", CH4_CONC = sc_1(CH4_AREA))
# Confirmed that the weirdly high fluxes in Passage 2, Week 1 were due to 
# using two different standard curves

@@

A set of `r n1` replicate ecosystems were generated
with 50 g of soil (sterile potting mix plus 5% inoculum by mass)
and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
These jars were initially kept at atmospheric concentrations of methane, but did not 
display significant methane oxidation rates after ~6 weeks. As a result, 
they were capped and spiked to
a headspace concentration of approximately 1000 ppm CH4. Both CH4 and CO2 were monitored
daily and it was determined that flushing and respiking was required every three
days in order to maintain aerobic conditions and an elevated CH4 concentration. After about 8 weeks
of incubation (2 weeks at 1000 ppm), we finally observed significant (negative)
methane fluxes. From the `r n1` jars, we selected the top 4 
(jars `r top_jars_pas1`) to homogenize and transfer to the next set of ecosystems.

For the first transfer, we used 10% inoculum intead of 5% inoculum and generated
`r n2` jars. After one week we observed
significant, negative methane fluxes that exceeded the first set of ecosystems. This
may have been a response to selection, a biomass effect due to the larger inoculum,
or both. After two weeks, we observed dampened fluxes compared to week one, but still
greater flux than before the first transfer. The elevated fluxes in week 1 compared
to week 2 could have been a result of using two different standard curves for t0 and t1
and for t2 and t3 in week 1 or due to some temporal effect during the incubations 
(e.g. response to water addition or initial available nutrients or growth).
In order to truly demonstrate a response to selection we will generate at
least one more set of ecosystems through a transfer with the same inoculum size
(10%) using jars `r top_jars_pas2` from passage 2.

@@

conc$selected[conc$JAR %in% top_jars_pas1 & conc$transfer == '1'] <- T
conc$selected[conc$JAR %in% top_jars_pas2 & conc$transfer == '2'] <- T
conc$selected[!(conc$JAR %in% top_jars_pas1) & conc$transfer == '1'] <- F
conc$selected[!(conc$JAR %in% top_jars_pas2) & conc$transfer == '2'] <- F
```

@@

flux <- 
flux %>%
  mutate(tidy_lm = map(fit_lm, tidy),
         tidy_glm = map(fit_glm, tidy)) %>% 
unnest(tidy_lm, tidy_glm) %>% 
filter(term == 'days') %>% 
select(transfer, Date, JAR, selected, lm_umold = estimate, glm_umold = estimate1)  %>% 
mutate(glm_umold_exp = exp(glm_umold))

all_par <- flux$lm_umold[flux$Date == '20190527']
offspr1 <- flux$lm_umold[flux$Date == '20190607']
offspr2 <- flux$lm_umold[flux$Date == '20190614']
sel_par1 <- flux$lm_umold[flux$transfer == '1' & flux$selected == T]
sel_par2 <- flux$lm_umold[flux$Date == '20190614' & flux$selected == T]

all_par_log <- flux$glm_umold[flux$Date == '20190527']
offspr1_log <- flux$glm_umold[flux$Date == '20190607']
offspr2_log <- flux$glm_umold[flux$Date == '20190614']
sel_par1_log <- flux$glm_umold[flux$transfer == '1' & flux$selected == T]
sel_par2_log <- flux$glm_umold[flux$Date == '20190614' & flux$selected == T]

all_par_exp <- flux$glm_umold_exp[flux$Date == '20190527']
offspr1_exp <- flux$glm_umold_exp[flux$Date == '20190607']
offspr2_exp <- flux$glm_umold_exp[flux$Date == '20190614']
sel_par1_exp <- flux$glm_umold_exp[flux$transfer == '1' & flux$selected == T]
sel_par2_exp <- flux$glm_umold_exp[flux$Date == '20190614' & flux$selected == T]
```

@@

ggplot(flux, aes(x = lm_umold)) +
  geom_histogram(aes(fill = interaction(selected, Date))) +
  facet_grid(Date + transfer ~ .) +
  scale_fill_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) +
  geom_vline(xintercept = mean(all_par), color = 'darkred') +
  geom_vline(xintercept = mean(offspr1), color = 'darkblue') +
  geom_vline(xintercept = mean(offspr2), color = 'darkorange') +
  geom_vline(xintercept = mean(sel_par1), color = 'red') +
  geom_vline(xintercept = mean(sel_par2), color = 'orange') +
  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "LM Fluxes")

ggplot(flux, aes(x = Date, y = lm_umold)) +
  facet_wrap(~ transfer + Date, scales = 'free_x') + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = interaction(selected, Date))) +
  scale_color_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) + 
  geom_hline(yintercept = mean(all_par), color = 'darkred') +
  geom_hline(yintercept = mean(offspr1), color = 'darkblue') +
  geom_hline(yintercept = mean(offspr2), color = 'darkorange') +
  geom_hline(yintercept = mean(sel_par1), color = 'red')  + 
  geom_hline(yintercept = mean(sel_par2), color = 'orange')  + 
  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "LM Fluxes")

ggplot(flux, aes(x = glm_umold)) +
  geom_histogram(aes(fill = interaction(selected, Date))) +
  facet_grid(Date + transfer ~ .) +
  scale_fill_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) +
  geom_vline(xintercept = mean(all_par_log), color = 'darkred') +
  geom_vline(xintercept = mean(offspr1_log), color = 'darkblue') +
  geom_vline(xintercept = mean(offspr2_log), color = 'darkorange') +
  geom_vline(xintercept = mean(sel_par1_log), color = 'red') +
  geom_vline(xintercept = mean(sel_par2_log), color = 'orange') +
  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "GLM Fluxes")

ggplot(flux, aes(x = Date, y = glm_umold)) +
  facet_wrap(~ transfer + Date, scales = 'free_x') + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = interaction(selected, Date))) +
  scale_color_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) + 
  geom_hline(yintercept = mean(all_par_log), color = 'darkred') +
  geom_hline(yintercept = mean(offspr1_log), color = 'darkblue') +
  geom_hline(yintercept = mean(offspr2_log), color = 'darkorange') +
  geom_hline(yintercept = mean(sel_par1_log), color = 'red') +
  geom_hline(yintercept = mean(sel_par2_log), color = 'orange') +
  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "GLM Fluxes")

ggplot(flux, aes(x = glm_umold_exp)) +
  geom_histogram(aes(fill = interaction(selected, Date))) +
  facet_grid(Date + transfer ~ .) +
  scale_fill_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) +
  geom_vline(xintercept = mean(all_par_exp), color = 'darkred') +
  geom_vline(xintercept = mean(offspr1_exp), color = 'darkblue') +
  geom_vline(xintercept = mean(offspr2_exp), color = 'darkorange') +
  geom_vline(xintercept = mean(sel_par1_exp), color = 'red') +
  geom_vline(xintercept = mean(sel_par2_exp), color = 'orange') +
  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "GLM Fluxes")

ggplot(flux, aes(x = Date, y = glm_umold_exp)) +
  facet_wrap(~ transfer + Date, scales = 'free_x') + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = interaction(selected, Date))) +
  scale_color_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) + 
  geom_hline(yintercept = mean(all_par_exp), color = 'darkred') +
  geom_hline(yintercept = mean(offspr1_exp), color = 'darkblue') +
  geom_hline(yintercept = mean(offspr2_exp), color = 'darkorange') +
  geom_hline(yintercept = mean(sel_par1_exp), color = 'red') +
  geom_hline(yintercept = mean(sel_par2_exp), color = 'orange') +
  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "GLM Fluxes")
```

@@

nsh1 <- ns_heritability(all_par, offspr1, sel_par1) 
nsh2 <- ns_heritability(all_par, offspr2, sel_par1) 
nsh1_log <- ns_heritability(all_par_log, offspr1_log, sel_par1_log)
nsh2_log <- ns_heritability(all_par_log, offspr2_log, sel_par1_log)
nsh1_exp <- ns_heritability(all_par_exp, offspr1_exp, sel_par1_exp)
nsh2_exp <- ns_heritability(all_par_exp, offspr2_exp, sel_par1_exp)
```

Based on the mean ecosystem-level trait values from the first and second set
of ecosystems (week 2), we calculated narrow-sense heritability as $h^2 = R / S$, which
was `r myround(nsh2, 2)` for the linear model and `r myround(nsh2_log, 2)` for the exponential
model and `r myround(nsh2_exp, 2)` for the back-transformed exponential coefficient. 
Granted, this is also not a valid estimate, because of the difference
in inoculum size between transfer 1 and 2. We will calculate this again between
transfer 2 and 3 to get a more accurate estimate.

@@

```{r identify_top_jars} 
flux %>% 
  group_by(Date) %>% 
  top_n(-4, lm_umold) %>% 
  kable(caption = "Linear model fluxes. transfer is the ecosystem generation,
        Date is the flux measurement date, JAR lists the top 4 jars from that sample
        date in terms of methane flux from a linear model, 
        selected indicates whether they were the
        selected parents for transfer 1, lm_umold is uMol/d flux based on a linear
        regression and glm_umold is uMol/d flux based on a glm with family = Gaussian
        and link = log") %>% 
  kable_styling()
flux %>% 
  group_by(Date) %>% 
  top_n(-4, glm_umold) %>% 
  kable(caption = "Generalized linear model link = log fluxes. Same as above except
        JAR indicates top 4 jars from a glm.") %>% 
  kable_styling()
flux %>% 
  group_by(Date) %>% 
  top_n(-4, glm_umold_exp) %>% 
  kable(caption = "Generalized linear model link = log fluxes. Same as above except
        JAR indicates top 4 jars from a glm and instead of slopes it's the exp(coef)
        so it's a multiplicative factor.") %>% 
  kable_styling()
```
